---
title: "52414 Lab Encryption Guide Revised v2"
author: "52414"
date: "5/09/2018"
output: html_document
---
A short guide on file encryption using public key and `sodium`  

First we need to install/load the necessary Libraries:  
Sodium is in CRan, while the wrapper cypher must be installed from github
```{r}
library("sodium")
library("cyphr")
library("repmis")
# To install cypher:
# remotes::install_github("ropensci/cyphr", upgrade = FALSE)
```


Loading the key object from github:  
Go to [Public_Key](https://github.com/DataScienceHU/DataAnalysisR/blob/master/guides/pub_cyphr_52414?raw=true) and the public key should be downloaded to your computer. 

```{r}
path.to.public.key <- "fill in your path to downloaded public key here"
load(path.to.public.key) #loading the downloaded public key into R
```


Encrypt your lab rmd file:
```{r}
rmd.path.name.input <- "fill in your path to rmd file  you want to encrypt"
rmd.path.name.output <- "fill in the path to the wanted encrypted ouput name"
cyphr::encrypt_file(rmd.path.name.input, pub_cyphr , rmd.path.name.output)
```

Now please upload your encypted RMD file to github:
https://github.com/DataScienceHU/DataAnalysisR/blob/master/Lab1/Lab1%20submissions/
