---
title: "52414 Lab Encryption Guide"
author: "52414"
date: "4/25/2018"
output: html_document
---
A short guide on file encryption using public key and `sodium`  

First you need to install/load the necessary Libraries.  
Both `Sodium` and `repmis` are in CRAN (and thus can be installed via `install.packages()` ).
However, `cypher` must be installed from github using the command that starts with `remotes::`
```{r}
library("sodium")
library("cyphr")
library("repmis") #see line 17
# To install cypher:
# remotes::install_github("ropensci/cyphr", upgrade = FALSE)
```

In order to encrypt a file, you must use the public-key stored in our github repository.  
Load the public `key` object from github and store the object in memory using the following command:
```{r}
source_data("https://github.com/DataScienceHU/DataAnalysisR/blob/master/guides/public_key_52414.RData?raw=True")
```

Now you are ready to encrypt your lab1 rmd file.  
Specify the path of your rmd file and the path you want to save your encripted rmd file in the following script:

```{r}
print(getwd()) #this can help you locate your rmd file
rmd.path.name.input <- "add your path to your rmd file here"
rmd.path.name.output <- "add the wanted path to the new encrypted file here"
# Below is the actual encryption command:
cyphr::encrypt_file(rmd.path.name.input, public.key ,rmd.path.name.output)
```

Last step: Upload your encypted RMD file to github in the following link:
`https://github.com/DataScienceHU/DataAnalysisR/tree/master/Lab1/Lab1_submissions/`
